---
title: "E-cig craving project"
output: html_document
author: Haiwen Lei & Xiaohan Zhang
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```


### load the data set 
```{r}
load("D:/DATA SCIENCE PROJECT/E-cig-craving-Project/data/36498-1001-Data.rda")
load("D:/DATA SCIENCE PROJECT/E-cig-craving-Project/data/36498-1002-Data.rda")
load("D:/DATA SCIENCE PROJECT/E-cig-craving-Project/data/36498-2001-Data.rda")
load("D:/DATA SCIENCE PROJECT/E-cig-craving-Project/data/36498-3001-Data.rda")
```


```{r}
library(tidyverse)
library(ggplot2)
wave1_adult <- da36498.1001
wave1_youth_parent <- da36498.1002


#select variables and data cleaning, the variables are chosen based on the group's knowledge of the subject. 
#We looked into the code book and questionnaire for the vairable selection. 

chosen_variable_wave1<- wave1_adult %>%
  select( 
         R01_AC1002, ##Have you ever smoked a cigarette, even one or two puffs?
         R01_AC1003, ##Do you now smoke cigarettes . . .
         R01_AC1004, ##In the past 30 days, have you smoked a cigarette, even one or two puffs?
         R01_AC9050, ##How harmful do you think cigarettes are to health?
         R01_AE1001, ## Have you ever seen or heard of an electronic cigarette or e-cigarette before this study?
         R01_AE1002, ##Have you ever used an e-cigarette, such as NJOY, Blu, or Smoking Everywhere, even one or two times?
         R01_AE1003, ##Do you now use e-cigarettes . . .
         R01_AE1004, ##In the past 30 days, have you used an e-cigarette, even one or two times?
         R01_AE1022, ##On how many of the past 30 days did you use [e-cigarette / e-cigarette cartridges]?
         R01_AE1008, ##When you first started using e-cigarettes, did you use cartridges or disposable e-cigarettes flavored to taste like menthol, mint, clove, spice, candy, fruit, chocolate, alcohol (such as wine or cognac), or other sweets?
         R01_AE1090, ##Do you own an e-cigarette?
         R01_AE1099, ##Is using e-cigarettes less harmful, about the same, or more harmful than smoking cigarettes?
         R01_AE1100, ##Have you ever used e-cigarettes fairly regularly?
         R01R_A_RACECAT3_IMP, ##Race category
         R01R_A_BMI, ## BMI 
         R01R_A_SEXORIENT2,## Sexual Orientation
         R01R_A_AC1006, ##Age range when first smoked part or all of a cigarette
         R01R_A_AE1006, ##Age range when first time used an e-cigarette, even one or two times
         R01R_A_AE1007, ##Age range when first started using e-cigarettes fairly regularly
         R01R_A_AE1020, ## Age range when first started using e-cigarettes every day
         R01R_A_AM0018, ## Highest grade or level of school completed
         R01R_A_AM0030, ## Total household income in the past 12 months
         R01_AN0015, ## Has strong cravings to use / smoke [tobacco products / specific product]
         R01_AN0025, ##Please rate your level of agreement for each statement using the following scale: 1= not true of me at all to 5=extremely true of me.  The following statements are about the use of tobacco and nicotine in general.  I frequently crave [ND FILL2].
         R01_AN0025E, ##Please rate your level of agreement for each statement using the following scale: 1= not true of me at all to 5=extremely true of me.  I frequently crave e-cigarettes.
         R01_AN0215, ## Thinking back to [the time you tried to quit / the last time you tried to quit / when you quit] in the past 12 months, did you use any of the following to help you: counseling, a telephone help line or quit line, books, pamphlets, videos, a quit tobacco clinic, class, or support group, or an internet or web-based program?
         R01_AN0215E, ## Thinking back to [the time you tried to quit / the last time you tried to quit / when you quit] using e-cigarettes in the past 12 months, did you use any of the following to help you: counseling, a telephone help line or quit line, books, pamphlets, videos, a quit tobacco clinic, class, or support group, or an internet or web-based program?
         R01_AN0250, ## To what extent, if at all, do you believe nicotine to be the main substance in tobacco that makes people want to use tobacco products?
         R01_AR0110, ## In the past 30 days, how often, if at all, have you noticed the health warnings on packages of e-cigarettes?
         R01_AR0142, ## In the past 12 months, how often have you thought about the chemicals contained in tobacco products?
         R01_AX0005, ## How would you describe your overall opinion of using tobacco?
         R01_AX0090, ##In general, how would you rate your physical health?
         R01_AX0091, ##In general, how would you rate your mental health, which includes stress, depression, and problems with emotions?
         R01_AX0004,## In the past 30 days, how often did you think about the harm your tobacco use might be doing to you?
         R01_AX0064_01, ## Work setting: Indoors Is your main work area indoors, outdoors, or in a vehicle? Choose all that apply
         R01_AX0064_02, ## Work setting: Outdoors Is your main work area indoors, outdoors, or in a vehicle? Choose all that apply
         R01_AX0064_03, ## Work setting: Vehicle Is your main work area indoors, outdoors, or in a vehicle? Choose all that apply
         R01_AX0065, ## How recently someone smoked around you while at work
         R01_AR0144, ## Lived with a regular smoker who smoked inside your home during childhood
         R01_AX0104,## Extent to which using / past use of tobacco products damaged your health
         R01_AX0105, ## Extent to which you are worried that using / your past use of tobacco products will damage your health in the future
         R01R_A_RACECAT3,## Race category
         R01R_A_CUR_ESTD_CIGS, ## Has ever smoked a cigarette, has smoked more than 100 cigarettes in lifetime, and currently smokes every day or some days.
         R01R_A_EDY_CIGS, ## Has ever used an cigarette, has used fairly regularly, and uses every day.
         R01R_A_SDY_CIGS, ## Has ever used an cigarette, has used fairly regularly, and uses some days
         R01R_A_P30D_CIGS, ## Has ever used an cigarette and has used at least once in the past 30 days
         R01R_A_CUR_ESTD_ECIG, ## Has ever used an e-cigarette, has used fairly regularly, and uses every day or some days.
         R01R_A_EDY_ECIG, ## Has ever used an e-cigarette, has used fairly regularly, and uses every day.
         R01R_A_SDY_ECIG, ## Has ever used an e-cigarette, has used fairly regularly, and uses some days
         R01R_A_P30D_ECIG, ## Has ever used an e-cigarette and has used at least once in the past 30 days
         R01R_A_AGECAT7_IMP, ## Imputed age range,
         )




######### Exploratory data analysis via data visualization



## race and e-cig smoking 
df1 <- chosen_variable_wave1 %>%
  mutate(R01_AN0025E = case_when(
    R01_AN0025E == "(1) 1 = Not true of me at all" ~ 1,
    R01_AN0025E == "(2) 2 = 2 on a scale of 1 to 5" ~ 2,
    R01_AN0025E == "(3) 3 = 3 on a scale of 1 to 5" ~ 3,
    R01_AN0025E == "(4) 4 = 4 on a scale of 1 to 5" ~ 4,
    R01_AN0025E == "(5) 5 = Extremely true of me" ~5
  ))%>%
   mutate(R01R_A_CUR_ESTD_ECIG = case_when(
    R01R_A_CUR_ESTD_ECIG == "(1) 1 = Yes" ~ 1,
    R01R_A_CUR_ESTD_ECIG == "(2) 2 = No" ~ 0
  ))%>%
  mutate(R01R_A_EDY_ECIG = case_when(
    R01R_A_EDY_ECIG == "(1) 1 = Yes" ~ 1,
    R01R_A_EDY_ECIG == "(2) 2 = No" ~ 0
  ))%>%
  mutate(R01R_A_SDY_ECIG = case_when(
    R01R_A_SDY_ECIG == "(1) 1 = Yes" ~ 1,
    R01R_A_SDY_ECIG == "(2) 2 = No" ~ 0
  ))%>%
  mutate(R01R_A_P30D_ECIG = case_when(
    R01R_A_P30D_ECIG == "(1) 1 = Yes" ~ 1,
    R01R_A_P30D_ECIG == "(2) 2 = No" ~ 0
  )) %>%
   mutate(R01R_A_CUR_ESTD_CIGS = if_else(R01R_A_CUR_ESTD_CIGS == "(1) 1 = Yes", 1,0 )) %>%
  mutate(R01R_A_EDY_CIGS = if_else(R01R_A_EDY_CIGS == "(1) 1 = Yes", 1,0 )) %>%
  mutate(R01R_A_SDY_CIGS = if_else(R01R_A_SDY_CIGS == "(1) 1 = Yes", 1,0 )) %>%
  mutate(R01R_A_P30D_CIGS = if_else(R01R_A_P30D_CIGS == "(1) 1 = Yes", 1,0 ))%>%
  rename(ecig_craving = R01_AN0025E) %>%
  rename(Race = R01R_A_RACECAT3) %>%
  rename(current_established_ecig_user = R01R_A_CUR_ESTD_ECIG) %>%
  rename(current_everyday_ecig_user = R01R_A_EDY_ECIG) %>%
  rename(current_someday_ecig_user = R01R_A_SDY_ECIG) %>%
  rename(current_past30day_ecig_user = R01R_A_P30D_ECIG)%>%
  rename(current_established_cig_user = R01R_A_CUR_ESTD_CIGS) %>%
  rename(current_everyday_cig_user = R01R_A_EDY_CIGS) %>%
  rename(current_someday_cig_user = R01R_A_SDY_CIGS) %>%
  rename(current_past30day_cig_user = R01R_A_P30D_CIGS)%>%
  rename(smooking_cigarettes = R01_AC1003)


   


#race distribution   
df1 %>%
  ggplot()+
  geom_bar(aes(x = Race))+
  labs(title = "race distribution")

df1 %>%
  ggplot(aes(Race)) +
  geom_bar(aes(y = (..count..)/sum(..count..)))+
  scale_y_continuous(labels=scales::percent) +
          ylab("relative frequencies")+
  labs(title = "race distribution")



df1 %>%       
  ggplot()+
  geom_col(aes(x = Race, y = current_established_ecig_user))+
  labs(title = "race distribution and current e-cig users")

## ecig craving and cig use 
df_graph3 <- df1%>%
  select(ecig_craving, smooking_cigarettes)%>%
  mutate(cig_usage = case_when(
    smooking_cigarettes == "(1) 1 = Every day"~ "Every day",
    smooking_cigarettes == "(2) 2 = Some days" ~ "Some days",
    smooking_cigarettes == "(3) 3 = Not at all" ~ "Not at all"
    ))%>%
  na.omit()%>%
  group_by(cig_usage, ecig_craving)%>%
  summarise(n = n())%>%
  mutate(freq = n / sum(n))%>%
  ungroup()
df_graph3%>%
  ggplot()+
  geom_col(aes(x = ecig_craving, y = freq, fill = cig_usage ))+
  labs(title = "ecig craving and cig use")



## smoking cigarettes and smoking e-cigarettes   
df1 %>%
  ggplot()+
  geom_col(aes(x = smooking_cigarettes, y = current_established_ecig_user))+
  labs(title = "smoking cigarettes and smoking e-cigarettes ")

df1 %>%
  ggplot()+
  geom_col(aes(x = smooking_cigarettes, y = current_everyday_ecig_user))+
  labs(title = "smoking cigarettes and smoking e-cigarettes ")


#Craving and established e-cig smokers
df1 %>%
  ggplot()+
  geom_col(aes(x =ecig_craving , y= current_established_ecig_user))+
  labs(title = "Craving and established e-cig smokers")

df1 %>%
  ggplot()+
  geom_bar(aes(x = ecig_craving))+
  labs(title = "Craving and established e-cig smokers") 




## worrying about tobacco damage and e-cig craving 
df_graph1 <- df1%>%
  select(R01_AX0004,ecig_craving)%>%
  mutate(frequent_worrying = case_when(R01_AX0004 == "(5) 5 = Very often" ~ "Very Often",
                                       R01_AX0004 == "(4) 4 = Often" ~ "Often",
                                       R01_AX0004 == "(3) 3 = Sometimes" ~ "Sometimes",
                                       R01_AX0004 == "(2) 2 = Rarely" ~ "Rarely",
                                       R01_AX0004 == "(1) 1 = Never" ~ "Never"
                                       ))%>%
  na.omit()%>%
  group_by(ecig_craving,frequent_worrying  )%>%
  summarise(n = n())%>%
  mutate(freq = n / sum(n))

df_graph1%>%
  ggplot(aes(x = ecig_craving, y = freq, color = frequent_worrying))+
  geom_point()+
  labs(title = "worrying about tobacco damage and e-cig craving" )


########################################################################

### appendix graphs 
df1 %>%       
  ggplot()+
  geom_col(aes(x = Race, y = current_established_ecig_user))


df1 %>% 
  ggplot()+
  geom_col(aes(x = Race, y = current_everyday_ecig_user))

df1 %>% 
  ggplot()+
  geom_col(aes(x = Race, y = current_someday_ecig_user))

df1 %>%   
  ggplot()+
  geom_col(aes(x = Race, y = current_past30day_ecig_user))


df1 %>%
  ggplot()+
  geom_col(aes(x = current_established_ecig_user, y = current_established_cig_user))

df1 %>%
  ggplot()+
  geom_col(aes(x = current_everyday_ecig_user, y = current_everyday_cig_user))

df1 %>%
  ggplot()+
  geom_col(aes(x = current_someday_ecig_user, y = current_someday_cig_user))
```

### Decision tree 
The decision tree is based on the 50 chosen variables with our subject of knowledge. 
```{r}
library(vip)
library(patchwork)
library(tidyverse)
library(tidymodels)
library(rpart)
library(rpart.plot)
library(parsnip)
library(workflows)
theme_set(theme_minimal())

## We dropped the observations that didn't answer the question for outcome variable(ecig_craving).  
df2<- chosen_variable_wave1%>%
  rename(ecig_craving = R01_AN0025E)%>%
  drop_na(ecig_craving)%>%
  rename(Race = R01R_A_RACECAT3) %>%
  rename(current_established_ecig_user = R01R_A_CUR_ESTD_ECIG) %>%
  rename(current_everyday_ecig_user = R01R_A_EDY_ECIG) %>%
  rename(current_someday_ecig_user = R01R_A_SDY_ECIG) %>%
  rename(current_past30day_ecig_user = R01R_A_P30D_ECIG)%>%
  rename(current_established_cig_user = R01R_A_CUR_ESTD_CIGS) %>%
  rename(current_everyday_cig_user = R01R_A_EDY_CIGS) %>%
  rename(current_someday_cig_user = R01R_A_SDY_CIGS) %>%
  rename(current_past30day_cig_user = R01R_A_P30D_CIGS)%>%
  rename(BMI = R01R_A_BMI )%>%
  rename(attitude = R01_AC9050)%>%
  rename(hhincome = R01R_A_AM0030)%>%
  rename(cig_crave = R01_AN0025)
  
df3<- wave1_adult%>%
  rename(ecig_craving = R01_AN0025E)%>%
  drop_na(ecig_craving)%>%
  rename(Race = R01R_A_RACECAT3) %>%
  rename(current_established_ecig_user = R01R_A_CUR_ESTD_ECIG) %>%
  rename(current_everyday_ecig_user = R01R_A_EDY_ECIG) %>%
  rename(current_someday_ecig_user = R01R_A_SDY_ECIG) %>%
  rename(current_past30day_ecig_user = R01R_A_P30D_ECIG)%>%
  rename(current_established_cig_user = R01R_A_CUR_ESTD_CIGS) %>%
  rename(current_everyday_cig_user = R01R_A_EDY_CIGS) %>%
  rename(current_someday_cig_user = R01R_A_SDY_CIGS) %>%
  rename(current_past30day_cig_user = R01R_A_P30D_CIGS)%>%
  rename(BMI = R01R_A_BMI )%>%
  rename(attitude = R01_AC9050)%>%
  rename(hhincome = R01R_A_AM0030)%>%
  rename(cig_crave = R01_AN0025)
 
set.seed(1234)

split <- initial_split(df2, 
                       prop = 0.7, 
                       strata = "ecig_craving")

train<- training(split)
test<- testing(split)



#creating a recipe for the supervised machine learning model
rec<- 
  recipe(ecig_craving~  ., data = df2) %>%
  step_dummy(all_nominal_predictors())


bake(prep(rec, train), new_data = train)

cart_mod <-decision_tree()  %>%
  set_engine(engine = "rpart") %>%
  set_mode(mode = "classification")

cart_wf <- workflow() %>%
  add_model(cart_mod)%>%
  add_recipe(rec)

cart_fit <- cart_wf %>%
  parsnip::fit( data = train)

y_predict <-predict(cart_fit, new_data = test) 


rpart.plot(x = cart_fit$fit$fit$fit,fallen.leaves=FALSE, tweak=1.2)


```
Decision tree have shown that being a everyday e-cigarette  and cigarette users tends to report e-cig craving. 
other variables deciding e-cig craving score include BMI, attitude towards smoking cigarette, wheter crave cigarette, and household income. 

### data imputation for missing values
We found many observations have missing values for prediction variables. We used the KNN machine learning method to impute the missing values. 
```{r}
library(tidymodels)
library(VIM)

df2_imp_f<- kNN(df2, k=5)
df2_imp_fct <- df2_imp_f%>%
  select(R01_AC1002: R01R_A_AGECAT7_IMP)


## converting to numeric 
df3[sapply(df3, is.factor)] <- lapply(df3[sapply(df3, is.factor)], as.numeric)
df2[sapply(df2, is.factor)] <- lapply(df2[sapply(df2, is.factor)], as.numeric)

df2_imputed<- kNN(df2, k=5)

df2_imp<- df2_imputed %>%
  select(R01_AC1002: R01R_A_AGECAT7_IMP)



```

### Feature selection
Three methods were used to select variables based on their importance. Before that, we conducted feature selection on original data frame to prove our knowledge of subject.Except for variables that are similar to e-cig craving score(we dropped them because they are not conducive for predicting ecig-craving), we found our chosen variables matches with feature selection. 
Then, we did feature selection on our imputed data frame.Three methods were used: random forest, machine learning algorithm, and boruta(another random forest method).After that, we targeted around 30 variables.    
```{r}

library(Boruta) #a feature ranking and selection algorithm based on random forests algorithm
library(caret)
library(party)


df3<- df3%>%
  select(-c(PERSONID,CASEID)) %>%
  mutate_all(~replace(., is.na(.), 0))

df2<- df2 %>%
  mutate_all(~replace(., is.na(.), 0))


#1. Random Forest Method
cf1 <- cforest(ecig_craving ~ . , data= df2, control=cforest_unbiased(mtry=2,ntree=50))
varimp(cf1)# get variable importance, based on mean decrease in accuracy

#on larger dataframe 
#cf2 <- cforest(ecig_craving ~ . , data= df3, control=cforest_unbiased(mtry=2,ntree=50))
#varimp(cf2)# get variable importance, based on mean decrease in accuracy

#on imputed dataframe 

cf3 <- cforest(ecig_craving ~ . , data= df2_imp, control=cforest_unbiased(mtry=2,ntree=50))
varimp(cf3)# get variable importance, based on mean decrease in accuracy


#2. Variable Importance from Machine Learning Algorithms
set.seed(11)
rPartMod <- train(ecig_craving ~ ., df2, method="rpart")
rpartImp <- varImp(rPartMod)
print(rpartImp)

set.seed(111)
rPartMod2 <- train(ecig_craving ~ ., df3, method="rpart")
rpartImp2 <- varImp(rPartMod2)
print(rpartImp2)
#deleting 
#R01_AN0090E	
#R01_AN0010E				
#R01_AN0085E				
#R01_AN0055E				
#R01_AN0050E				
#R01_AN0020E				
#R01_AN0065E different ways of asking about e-cig craving 


df3_d<- df3%>%
  select(-c(R01_AN0090E,			
R01_AN0010E,			
R01_AN0085E,			
R01_AN0055E,			
R01_AN0050E,			
R01_AN0020E,			
R01_AN0065E,
R01_AN0060E,		
R01_AN0095E,			
R01_AN0100E,			
R01_AN0045E,			
R01_AN0030E,
R01_AN0015E))

set.seed(222)
rPartMod3 <- train(ecig_craving ~ ., df3, method="rpart")
rpartImp3 <- varImp(rPartMod3)
print(rpartImp3)

## imputed dataframe  
set.seed(333)
rPartMod4 <- train(ecig_craving ~ ., df2_imp, method="rpart")
rpartImp4 <- varImp(rPartMod4)
print(rpartImp4)


# 3.Boruta
set.seed(11)
boruta_output <- Boruta(ecig_craving ~ ., data=na.omit(df2), doTrace=2)
names(boruta_output)
boruta_signif <- getSelectedAttributes(boruta_output, withTentative = TRUE)
print(boruta_signif)

set.seed(22)
boruta_output2 <- Boruta(ecig_craving ~ ., data=na.omit(df2_imp), doTrace=2)
boruta_signif2 <- getSelectedAttributes(boruta_output2, withTentative = TRUE)
print(boruta_signif2)


imps <- attStats(boruta_output2)
imps2 = imps[imps$decision != 'Rejected', c('meanImp', 'decision')]
head(imps2[order(-imps2$meanImp), ])  # descending sort

plot(boruta_output2, cex.axis=.7, las=2, xlab="", main="Variable Importance")




```


### Supervised ML model with imputed data
We constructed supervised ML model with imputed data and used the models to predict e-cig craving.
Two models were used random forest and lasso regression.We used the factor data frame for Random forest with classification mode. We also converted the factor df to numeric df for regression models(second random forest and lasso regression).   
```{r}
library(vip)
library(patchwork)
library(tidyverse)
library(tidymodels)
library(rpart)
library(rpart.plot)
library(parsnip)
library(workflows)
library(ranger)
library(yardstick)

theme_set(theme_minimal())

df_dt<- df2_imp_fct %>%
  select(R01_AC1003,
  attitude,
  R01_AE1003,
  R01_AE1004,
  R01_AE1022,
  R01_AE1090,
  R01_AE1100,
  R01R_A_AE1006,
  R01R_A_AE1007,
  R01R_A_AE1020,
  R01_AN0015,
  cig_crave,
  R01_AN0215,
  R01_AN0215E,
  R01_AR0110,
  R01_AX0005,
  R01_AX0091,
  R01_AX0004,
  R01_AR0144,
  R01_AX0104,
  R01_AX0105,
  current_established_cig_user,
  current_everyday_cig_user,
  current_someday_cig_user,
  current_past30day_cig_user,
  current_established_ecig_user,
  current_everyday_ecig_user,
  current_someday_ecig_user,
  current_past30day_ecig_user,
  R01R_A_AGECAT7_IMP,
  hhincome,
  R01R_A_SEXORIENT2,
  R01_AX0064_01,
  ecig_craving)


df_dt2<- df2_imp %>%
  select(R01_AC1003,
  attitude,
  R01_AE1003,
  R01_AE1004,
  R01_AE1022,
  R01_AE1090,
  R01_AE1100,
  R01R_A_AE1006,
  R01R_A_AE1007,
  R01R_A_AE1020,
  R01_AN0015,
  cig_crave,
  R01_AN0215,
  R01_AN0215E,
  R01_AR0110,
  R01_AX0005,
  R01_AX0091,
  R01_AX0004,
  R01_AR0144,
  R01_AX0104,
  R01_AX0105,
  current_established_cig_user,
  current_everyday_cig_user,
  current_someday_cig_user,
  current_past30day_cig_user,
  current_established_ecig_user,
  current_everyday_ecig_user,
  current_someday_ecig_user,
  current_past30day_ecig_user,
  R01R_A_AGECAT7_IMP,
  hhincome,
  R01R_A_SEXORIENT2,
  R01_AX0064_01,
  ecig_craving)


set.seed(12345)
split2 <- initial_split(df_dt, 
                       prop = 0.7, 
                       strata = "ecig_craving")

train2<- training(split2)
test2<- testing(split2)



#creating a recipe for the supervised machine learning model



rec2<- 
  recipe(ecig_craving~  ., data = df_dt) %>%
  step_dummy(all_nominal_predictors())
  

folds <- vfold_cv(data = train2, v = 10, repeats = 1)

bake(prep(rec2, train2), new_data = train2)


# Random forest classification 

rand_fore <-rand_forest(trees = 1000) %>% 
  set_engine("ranger") %>% 
  set_mode("classification")

rand_wf <- workflow() %>%
  add_recipe(rec2) %>%
  add_model(rand_fore) 
 
rand_cv <- rand_wf %>%
 tune_grid(grid = 10,
            control = control_grid(save_pred = TRUE),
           resamples = folds)


rand_rmse<- rand_cv %>%
  collect_metrics(summarize = TRUE)

rand_rmse %>%
  ggplot(aes(x= .config, y= mean))+
  geom_point()

rand_fit <- rand_wf%>%
  parsnip::fit(train2)

best_model_rmse <- bind_cols(
  test2,
  predict(object = rand_fit, new_data = test2)
) 



yardstick::accuracy(data =best_model_rmse,
                    truth = ecig_craving,
                    estimate = .pred_class)

yardstick::precision(data = best_model_rmse,
         truth = ecig_craving,
         estimate = .pred_class)

yardstick::sensitivity(data = best_model_rmse,
         truth = ecig_craving,
         estimate = .pred_class)

yardstick::recall(data = best_model_rmse,
         truth = ecig_craving,
         estimate = .pred_class)

conf_mat(data = best_model_rmse,
         truth = ecig_craving,
         estimate = .pred_class)

## converting to numeric 

 best_model_rmse[sapply(best_model_rmse, is.factor)] <- lapply(best_model_rmse[sapply(best_model_rmse, is.factor)],as.numeric)
 
 rmse(data = best_model_rmse, truth = ecig_craving, estimate = .pred_class )


## regression model  

set.seed(123456)
split3 <- initial_split(df_dt2, 
                       prop = 0.7, 
                       strata = "ecig_craving")

train3<- training(split3)
test3<- testing(split3)


rec3<- 
  recipe(ecig_craving~  ., data = df_dt2) %>%
  step_dummy(all_nominal_predictors())
  

folds2 <- vfold_cv(data = train3, v = 10, repeats = 1)

############### randome forest

rand_fore2 <-rand_forest(trees = 1000) %>% 
  set_engine("ranger") %>% 
  set_mode("regression")

rand_wf2 <- workflow() %>%
  add_recipe(rec3) %>%
  add_model(rand_fore2) 
 
rand_cv2 <- rand_wf2 %>%
 tune_grid(grid = 10,
            control = control_grid(save_pred = TRUE),
           metrics = metric_set(mae, rmse),
           resamples = folds2)


rand_rmse2<- rand_cv2 %>%
  collect_metrics(summarize = TRUE)

rand_rmse2 %>%
   ggplot(aes(x= .config, y= mean, color = .metric))+
  geom_point()

rand_fit2 <- rand_wf2%>%
  parsnip::fit(train3)

best_model_rmse2 <- bind_cols(
  test3,
  predict(object = rand_fit2, new_data = test3)
) 


rmse(data = best_model_rmse2, truth = ecig_craving, estimate = .pred)
mae(data = best_model_rmse2, truth = ecig_craving, estimate = .pred)

########### lasso 
# lasso

lasso_grid <- grid_regular(penalty(), levels = 10)

lasso_mod <- linear_reg(
  penalty = tune(), 
  mixture = 1
) %>%
  set_engine("glmnet")

lasso_wf <- workflow() %>%
  add_recipe(rec3) %>%
  add_model(lasso_mod) 

lasso_cv <- lasso_wf %>%
  tune_grid(
    resamples = folds2,
    grid = lasso_grid,
    metrics = metric_set(mae, rmse)
  )

lasso_rmse<- lasso_cv %>% 
  collect_metrics(summarize = TRUE)

lasso_cv %>% 
  show_best(metric = "rmse", n = 1)

lasso_best <- lasso_cv %>%
  select_best(metric = "rmse")

lasso_final <- finalize_workflow(
  lasso_wf,
  parameters = lasso_best
)

lasso_fit<- lasso_final%>%
  parsnip::fit(data = train3)

lasso_coefs <- lasso_final %>%
  parsnip::fit(data = train3) %>%
  extract_fit_parsnip() %>%
  vi(lambda = lasso_best$penalty) 

print(lasso_coefs) #this tells us the coefficients of each variable and their magnitude

lasso_rmse %>%
  ggplot(aes(x= .config, y= mean, color = .metric))+
  geom_point()

best_model_rmse3 <- bind_cols(
  test3,
  predict(object = lasso_fit, new_data = test3)
) 

rmse(data = best_model_rmse3, truth = ecig_craving, estimate = .pred)
mae(data = best_model_rmse3, truth = ecig_craving, estimate = .pred)


```
lasso and random forest model have similar rmse about 0.87 and 0.88. 


### Extension- cluster analysis
clusters among the people who crave e-cig. 
```{r}
library(factoextra)
library(patchwork)







set.seed(20220412)

# total within sum of squares
fviz_nbclust(df_dt2, FUN = kmeans, method = "wss")

# total silhouette width
fviz_nbclust(df_dt2, FUN = kmeans, method = "silhouette")

# gap statistic
fviz_nbclust(df_dt2, FUN = kmeans, method = "gap_stat")

## 2 or 8 clusters 

#### creating PCA
pca<- prcomp(df_dt2)
summary(pca)

pcs <- pca %>%
.$x %>%
  as_tibble()
######

pcs <-bind_cols(
  select(pcs, PC1, PC2),
  df_dt2)

  kmeans<- kmeans(
   df_dt2,
   centers = 2)
  
 cluster<- bind_cols(
    pcs,
    cluster = kmeans$cluster
  )
  
  ggplot() +
  geom_point(
    data = cluster,
     mapping = aes(x = PC1, y = PC2, color = factor(cluster))) +
    labs(
      title = "K-Means with 2 and PCA") 

# 8 clusters  
  
pca<- prcomp(df_dt2)
summary(pca)

pcs <- pca %>%
.$x %>%
  as_tibble()  
  
pcs2 <-bind_cols(
  select(pcs, PC1, PC2, PC3, PC4, PC5, PC6, PC7, PC8),
  df_dt2)

  kmeans2<- kmeans(
   df_dt2,
   centers = 8)
  
 cluster2 <- bind_cols(
    pcs,
    cluster = kmeans2$cluster
  )
  
  ggplot() +
  geom_point(
    data = cluster2,
     mapping = aes(x = PC1, y = PC2, color = factor(cluster))) +
    labs(
      title = "K-Means with 2 and PCA") 

```




